---
title: "Starwars Network Analysis"
author: "Rafaela Becerra"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Packages needed
```{r}
library(igraph)
```
## Load the Star Wars network created.
```{r}
setwd("C:/Users/Rafaela Becerra/Desktop/UC3M/Network_Analysis/Assignments/Starwars")
sw.network <- read_graph('sw.net.txt', format='gml')
```

## Order and size of the network.

```{r}
go<- gorder(sw.network)
print(paste('The order is: ',go))
```
```{r}
gs<- gsize(sw.network)
print(paste('The size is: ',gs))
```


## Induced subgraph that includes the following characters:

- DARTH VADER (number 1 of the list of vertices),
- R2-D2 (2),
- CHEWBACCA (3),
- OBI-WAN (9),
- EMPEROR (13),
- C-3PO (25),
- JABBA (30), 
- GREEDO (31),
- YODA (35), 
- BOBA FETT (49), 
- BERU (52), 
- CAPTAIN ANTILLES (66),
- LUKE (67),
- CAMIE (68), 
- BIGGS (69), 
- LEIA (70), 
- MOTTI (71),
- TARKIN (72), 
- HAN (73), 
- DODONNA (74), 
- GOLD LEADER (75), 
- WEDGE (76),
- RED LEADER (77), 
- RED TEN (78), 
- GOLD FIVE (79), 
- RIEEKAN (80),
- DERLIN (81), 
- ZEV (82), 
- PIETT (83),
- OZZEL (84),
- DACK (85), 
- JANSON (86), 
- NEEDA (87), 
- LANDO (88), 
- JERJERROD (89), 
- BIB FORTUNA (90),
- BOUSHH (91), and
- ADMIRAL ACKBAR (92).

```{r}
include_v <- c(1,2,3,9,13,25,30,31,35,49,52,66, 67, 68,69, 70, 71:92)
induced.1 <- induced_subgraph(sw.network, include_v)
induced.1
```

## Order and size of the induced subgraph.
```{r}
go<- gorder(induced.1)
print(paste('The order is: ',go))
```
```{r}
gs<- gsize(induced.1)
print(paste('The size is: ',gs))
```

## Basic visual representation of such subgraph.
```{r}
vertex_attr_names(induced.1)
edge_attr(induced.1)
```
```{r}
la <- layout_in_circle(induced.1)
#Graph display of the network.
plot.igraph(induced.1,
            vertex.label=V(induced.1)$name,
            layout=la, 
            edge.color="lightblue",
            vertex.size=5,
            vertex.label.dist=1,
            edge.width=E(induced.1)$weight/10,
            vertex.label.color="black",
            vertex.label.cex=0.5,
            vertex.color="blue",
            vertex.size=15)
```

By this graph we can see a graphical representation of a prespecified subset of the vertices of the network. In this case, we have determined the width of the edges as the weights or the number of scenes in which the characters of the vertices appear together. As display, Gold Five appears alone in all of the episodes, conversely, Han with Chewbacca seems to be a repetitive duo on the scenes. Also, with major appearences together are Han with C-3PO, Luke or Leia, and other combinations like Luke and R2-D2, Leia and Chewbacca, or Leia and C-3PO.

Additionally, by the number of edges we can determine which characters appear with more characters, for example, we see that characters like Janson, Dack, Jerjerrod appear just with one other character and these appereances are not frequent as the width is really small.On the other side, characters like Darth Vader appear with several edges that connect it to other characters, so this means that it appears with much more characters than the others. 

## Induced subgraph composed by the other characters.
```{r}
ninclude_v <- subset(seq(1:111), !(seq(1:111) %in% include_v))
induced.2 <- induced_subgraph(sw.network, ninclude_v)
induced.2
```

```{r}
go<- gorder(induced.2)
print(paste('The order is: ',go))
```

```{r}
gs<- gsize(induced.2)
print(paste('The size is: ',gs))
```
```{r}
vertex_attr_names(induced.2)
edge_attr(induced.2)
```

```{r}
la <- layout_in_circle(induced.2)
#Graph display of the network.
plot.igraph(induced.2,
            vertex.label=V(induced.2)$name,
            layout=la, 
            edge.color="lightgreen",
            vertex.size=5,
            vertex.label.dist=1,
            edge.width=E(induced.2)$weight/5,
            vertex.label.color="black",
            vertex.label.cex=0.4,
            vertex.color="green",
            vertex.size=15)
```
We have repeated this graph for the second group of characters, as before, the width represents the number of scenes together. These characters seem to have less participation on the scenes because, in one side, the conexion between them is less as the number of edges is lower and second, the width is small, consequently we could said that this characters do not interact a lot with each other. Moreover, we can see that this group is separated in other groups, for checking this, we have changed the layout of the graph for a clearer visualization. 

```{r}
la <- layout_nicely(induced.2)
#Graph display of the network.
plot.igraph(induced.2,
            vertex.label=V(induced.2)$name,
            layout=la, 
            edge.color="lightgreen",
            vertex.size=5,
            vertex.label.dist=1,
            edge.width=E(induced.2)$weight/5,
            vertex.label.color="black",
            vertex.label.cex=0.4,
            vertex.color="green",
            vertex.size=15)
```
We can determine three groups in this subgraph which englobe one, characters that do not interact with other characters in the group at all, and two, groups that contain characters that interact within other characters in the same group but not with the other groups.

Moreover, the number of edges is not equal to the total, consequently, we are sure that there are interactions between this second subgraph and the first display in blue.


The difference between the total number of edges of the network minus the edges of each subgraph will be the number of edges joining the vertices between the two subgraphs.

```{r}
gs<- gsize(sw.network)-gsize(induced.2)-gsize(induced.1)
print(paste('The number of edges joining the vertices between the subgraphs is: ',gs))
```

```{r}
print('The number of neighbors or adjacent vertices of (1) is: ')
neighbors(sw.network,1)
```
```{r}
print('Their incident edges are: ')
incident(sw.network,1)
```
## Degree of DARTH VADER (1).
```{r}
print('The degree of (1) is: ')
degree(sw.network,1)
```
This is the highest if we compare it to the degree of the rest of the characters as we can check below in the degree sequence of the whole network. This means that this character appears with a lot of the other characters in the scenes.

## Degree sequence, the average degree and the degree distribution of the whole network without weights.

- The degree sequence:
```{r}
degree(sw.network)
#Check that it is twice the size graph
sum(degree(sw.network))
```

- The average degree:
```{r}
mean(degree(sw.network))
```

- Degree distribution

Frquency table
```{r}
table(degree(sw.network))
```

```{r}
hist(degree(sw.network,mode="total"),
     col="pink",
     main="Total degree distribution of the complete Starwars network",
     xlab="Degree",
     freq=FALSE)
```    

Degree distribution
```{r}
degree_distribution(sw.network)
```

- Plot of the distribution
```{r}
plot(0:max(degree(sw.network)),
     degree_distribution(sw.network),
     ylim=c(0,(max(degree_distribution(sw.network))+0.01)),
     col="pink",
     main="Degree distribution of complete Starwars network",
     xlab="Degree",
     ylab="Frequency",
     type="h")
points(0:max(degree(sw.network)),
       degree_distribution(sw.network),
       pch=20)
text(0:max(degree(sw.network)),
     degree_distribution(sw.network),
     labels=format(100*degree_distribution(sw.network), digits=2), cex=0.5, font=1, pos=3)
```
The degree corresponds to the number of edges incident on a vertex, which in this case corresponds to the number of characters among all which interacted in the scenes of the 1 to seven episodes of Starwars with this certain character represented by this vertex.

Now, we can see that Darth Vader appears with almost half of the characters in the scenes, consequently, is the character whith the highest degree. 

By looking at the frequency table of the degrees and the histogram, we see that most of the characters interact with at most 4 characters, and that there is one character that appears alone. 
The degree of character 1 Darth Vader is almost 7 times the average degree because this returned a value of 8 characters, this can mean that the degree sequence will show a skewed distribution to the right.

After, plotting the distribution of the degrees, which is constructed by the probabilities that a selected vertex in the network has a degree d in a range of degrees, we can see the skewed distribution, as most of the characters just had interactions in the scenes with fewer characters and the results with more than 10 characters are given by just one or two of them. 

## Degree sequence, the average degree and the degree distribution of the whole network with weights. 

- The degree sequence with weights:
```{r}
strength(sw.network)
#Check that it is twice the size graph
sum(strength(sw.network))
```

- The average degree with weights:
```{r}
mean(strength(sw.network))
```

- Degree distribution with weights

Frquency table with weights
```{r}
table(strength(sw.network))
```
```{r}
hist(strength(sw.network,mode="total"),
     col="purple",
     main="Total weighted degree distribution of the complete Starwars network",
     xlab="Degree",
     freq=FALSE)
```    
     
Degree distribution with weights
```{r}
graph_strength_distribution <- function (graph, cumulative = FALSE, ...)
{
  if (!is.igraph(graph)) {
    stop("Not a graph object")
  }
  cs <- graph.strength(graph, ...)
  hi <- hist(cs, -1:max(cs), plot = FALSE)$density
  if (!cumulative) {
    res <- hi
  }
  else {
    res <- rev(cumsum(rev(hi)))
  }
  res
}
graph_strength_distribution(sw.network)
```

- Plot of the distribution with weights
```{r}
plot(0:max(strength(sw.network)),
     graph_strength_distribution(sw.network),
     ylim=c(0,(max(graph_strength_distribution(sw.network))+0.01)),
     col="purple",
     main="Weighted degree distribution of complete Starwars network",
     xlab="Degree",
     ylab="Frequency",
     type="h")
points(0:max(strength(sw.network)),
       graph_strength_distribution(sw.network),
       pch=20)
```
If we consider the weights or the number of appearences that characters have with each other and not just the relations, we can see that the character with more participation is Han which has 360 scenes with other characters, next we have C-3PO with 310 and Darth Vader with 309. Consequently, even though Darth Vader is the character that appears with more characters, it is not the one that appears the most in these seven episodes. Moreover, the total number of weighted degrees is 4184, these means that we have 2092  connections if we consider the frequency. 

Additionally, the average weighted degree is approximately 38, which is much more than the unweighted degree which derives the importance of the frequency in this case. This number represents the number of scenes in average that a character appeared. If we check the histogram and the frequency of the weighted degrees, we can see that there are a lot of characters that have almost no appearences like Colonel Datto, Clone Commander, Tion Medon, etc. 

Because of this disparateness on the number of appearances the distribution is still skewed to the right.  


## Previous steps for the two induced subgraphs.

## Subgraph 1

## Degree
- The degree sequence:
```{r}
degree(induced.1)
#Check that it is twice the size graph
sum(degree(induced.1))
```

- The average degree:
```{r}
round(mean(degree(induced.1)))
```

- Degree distribution

Frquency table
```{r}
table(degree(induced.1))
```

```{r}
hist(degree(induced.1,mode="total"),
     col="lightblue",
     main="Total degree distribution of subgraph 1",
     xlab="Degree",
     freq=FALSE)
```    
Degree distribution
```{r}
degree_distribution(induced.1)
```

- Plot of the distribution
```{r}
plot(0:max(degree(induced.1)),
     degree_distribution(induced.1),
     ylim=c(0,(max(degree_distribution(induced.1))+0.01)),
     col="lightblue",
     main="Degree distribution of of subgraph 1",
     xlab="Degree",
     ylab="Frequency",
     type="h")
points(0:max(degree(induced.1)),
       degree_distribution(induced.1),
       pch=20)
text(0:max(degree(induced.1)),
     degree_distribution(induced.1),
     labels=format(100*degree_distribution(induced.1), digits=2), cex=0.5, font=1, pos=3)
```

If we take a look at the first subgroup, we can see that in this prespecified subset of the vertices lay 38 characters which create a network with 126 edges or the  connections among these characters and represent a total of 28.38%, if we consider the total edges on the whole network. 

This subgraph presents an average degree of ,approximately, 7 which is lower that what we got for the complete network (8), this can be attributed to the fact that there are more characters in the other group, consequently, the  connections between them are not taken into account. In fact, recall that the number of edges joinning both subgroups are 156.

If we check the frequency table of the degree distribution we see that the media is 2, and that the maximum number is 23, and that as the degree increases there are only ones for each number and a total of 7 characters that appear with more than 10 other characters. If we check the accumulated probability that a random vertex has less or equal to five degrees is more than 50%, so there is a high probability that, when just considering this group, the number of characters that act together is less than for the total network.

## Degree with weights

- The degree sequence with weights:
```{r}
strength(induced.1)
#Check that it is twice the size graph
sum(strength(induced.1))
```

- The average degree with weights:
```{r}
mean(strength(induced.1))
```

- Degree distribution with weights

Frquency table with weights
```{r}
table(strength(induced.1))
```
```{r}
hist(strength(induced.1,mode="total"),
     col="lightblue",
     main="Total weighted degree distribution of subgraph 1",
     xlab="Degree",
     freq=FALSE)
```    
     
Degree distribution with weights
```{r}
graph_strength_distribution(induced.1)
```

- Plot of the distribution with weights
```{r}
plot(0:max(strength(induced.1)),
     graph_strength_distribution(induced.1),
     ylim=c(0,(max(graph_strength_distribution(induced.1))+0.01)),
     col="lightblue",
     main="Weighted degree distribution of subgraph 1",
     xlab="Degree",
     ylab="Frequency",
     type="h")
points(0:max(strength(induced.1)),
       graph_strength_distribution(induced.1),
       pch=20)
```
Now, if we consider the weights, we see that the sum of the degrees is 2176, which is equivalent of having 1088 scenes between these characters, which represents 52% of the total, consequently, we could say that in this group are consider more characters that have performed a more interactive role, since in the overall we see that this is a smaller number of characters but the total scenes with other characters are very representative. 

Additionally, we see that the mean degree is around 57, 20 points higher that for the complete network, and that the distribution is more concetrated in small numbers, being the larger probability of 18% for 2 degrees and the second larger 10% for 4 degrees, but a skewness is generated because of other characters with major appearances. 

## Subgraph 2

## Degree
- The degree sequence:
```{r}
degree(induced.2)
#Check that it is twice the size graph
sum(degree(induced.2))
```

- The average degree:
```{r}
mean(degree(induced.2))
```

- Degree distribution

Frquency table
```{r}
table(degree(induced.2))
```

```{r}
hist(degree(induced.2,mode="total"),
     col="lightgreen",
     main="Total degree distribution of subgraph 2",
     xlab="Degree",
     freq=FALSE)
```    
Degree distribution
```{r}
degree_distribution(induced.2)
```

- Plot of the distribution
```{r}
plot(0:max(degree(induced.2)),
     degree_distribution(induced.2),
     ylim=c(0,(max(degree_distribution(induced.2))+0.01)),
     col="lightgreen",
     main="Degree distribution of of subgraph 2",
     xlab="Degree",
     ylab="Frequency",
     type="h")
points(0:max(degree(induced.2)),
       degree_distribution(induced.2),
       pch=20)
text(0:max(degree(induced.2)),
     degree_distribution(induced.2),
     labels=format(100*degree_distribution(induced.2), digits=2), cex=0.5, font=1, pos=3)
```

If we take a look at the second subgroup, we can see that in this prespecified subset of the vertices lay 73 characters which are more than 65% of the total characters, and that created an induced subgraph with 162 edges or the  connections among these characters and which represents a total of 36.49% if we consider the total edges on the whole network. 

This subgraph presents an averge degree of approximately 4, which is lower that what we got for the complete network (8), as in the case for the first subgroup, this can be attributed to the fact that the connections between the groups are not taken into account, and as we saw in the graphs in this subgroup lay a lot of characters that do not play a major role in these episodes. 

If we check the frequency table of the degree distribution we see that the media is 3, and that the maximum number is 26 which corresponds to Padme. Moreover, if we check the accumulated probability that a random vertex has less or equal to five degrees is more than 80%, so there is a higher probability when comparing to the other subgroup, reaffirming that in this group are more characters that did not participate much with other characters.

## Degree with weights

- The degree sequence with weights:
```{r}
strength(induced.2)
#Check that it is twice the size graph
sum(strength(induced.2))
```

- The average degree with weights:
```{r}
mean(strength(induced.2))
```

- Degree distribution with weights

Frquency table with weights
```{r}
table(strength(induced.2))
```
```{r}
hist(strength(induced.2,mode="total"),
     col="lightgreen",
     main="Total weighted degree distribution of subgraph 2",
     xlab="Degree",
     freq=FALSE)
```    
     
Degree distribution with weights
```{r}
graph_strength_distribution(induced.2)
```

- Plot of the distribution with weights
```{r}
plot(0:max(strength(induced.2)),
     graph_strength_distribution(induced.2),
     ylim=c(0,(max(graph_strength_distribution(induced.2))+0.01)),
     col="lightgreen",
     main="Weighted degree distribution of subgraph 2",
     xlab="Degree",
     ylab="Frequency",
     type="h")
points(0:max(strength(induced.2)),
       graph_strength_distribution(induced.2),
       pch=20)
```

Finally, if we take in consideration the weights or the number of scenes that the characters considered in this subgraph did together, we can see that the sum of degrees is just 832, which is equivalent to just 416 scenes together.  

Additionally, we see that the mean degree is around 11, lower than for the other subgraph and the complete network, consequently, we can see that in this group there are a lot of characters that participate less during the episodes if we take in account the number of scenes that did with other characters. If we check the probability that a randomly selected vertex has a degree considering weight equal or smaller than 5 is almost 48%, proving that in this group the interaction is representatively less than in the other case analysed.

#Check that the network is not a multigraph (Graphs with multiple directed edges
#sharing the same vertices)
```{r}
is_simple(sw.network)
```
## Random walk of length 5 in the network.
```{r}
random_walk(sw.network,start=1,steps=5,mode='all',stuck="return")
```
## Shortest circuit in the network.
```{r}
girth(sw.network,circle=TRUE)
```

The shortest circuit display involves 3 vertices which in this case correspond to the characters 'R2-D2', 'DARTH VADER' and 'CHEWBACCA'

## Check wether the network is connected or not.
```{r}
is_connected(sw.network)
```
Every vertex is not reachable from every other, consequently, the network is not connected and there will be characters that did not interact with others on the scenes of these episodes.

## Components of the network.
```{r}
components(sw.network)
```
By taking a look at the components we can see that there is one character Gold Five which did not do any scene with other characters, then we do not have a single component.

## Diameter, the associated path and the farthest vertices with no weights, as well as for their weighted versions.
```{r}
get_diameter(sw.network,weights=rep(1,gsize(sw.network)))
diameter(sw.network,weights=rep(1,gsize(sw.network)))
farthest_vertices(sw.network,weights=rep(1,gsize(sw.network)))
```


```{r}
get_diameter(sw.network)
diameter(sw.network)
farthest_vertices(sw.network)
```

The diameter will be the longest geodesic distance in the network, in this case if we do not take in account the weights or the ammount of scenes that the characters did together, we get that this corresponds to a diameter of 6, the path extracted shows the vertices involve correspond to:

1. DOFINE
2. NUTE GUNRAY
3. DARTH VADER  
4. CHEWBACCA     
5. KYLO REN      
6. GENERAL HUX  
7. COLONEL DATOO

Being the farthest vertices (1) and (7).

On the other side, if we take the diameter by weighting edges we get that this corresponds to 10, the path showed the characters:

1. RUNE         
2. DARTH MAUL   
3. GENERAL CEEL 
4. NUTE GUNRAY  
5. OBI-WAN      
6. LEIA         
7. BIGGS   
8. WEDGE        
9. JANSON 

Where the farthest vertices are (1) and (9)